Proving linearizability using forward simulations\thanks{An extended version of this paper including the missing proofs can be found at \cite{extended}.}

### 2017-05-05 draftCAV/queue.tex
Syntax of abstract state machines.

%\begin{figure} [t]
\begin{figure}[t]
\hspace{-10mm}
\begin{minipage}[t]{7.5cm}
\scriptsize{
{	%\centering
	\begin{align*}
%		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
		& \text{\underline{Shared State}} \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \{{\tt true},{\tt false}\} \\
%		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
%		& \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} &\mbox{ // the argument of an enqueue} \\
	\end{align*}
}}
\end{minipage}
\begin{minipage}[t]{7.5cm}
\begin{lstlisting}
void enq($\<Vals>$ x) {
  atomic rule inv(enq,x,k = new);
  atomic rule ret(enq,k);
}

$\<Vals>$ deq() {
  atomic rule inv(deq,k = new);
  atomic rule lin(deq,y,k);
  atomic rule ret(deq,y,k);
}
\end{lstlisting}
\end{minipage}

\begin{minipage}[t]{7.5cm}
\vspace{-3mm}
\begin{lstlisting}
rule inv(enq,x,k) = {
  $\mathsf{present}$(k) = true;
  $\mathsf{pending}$(k) = true;
  $\mathsf{arg}$(k) = x;
  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1) do
    $\mathsf{before}$(k1,k) = true;
}

rule lin(deq,y,k) = {
  choose a $\in$ {true, false};
  if ( $($a$\ \vee\ \forall k'.\ \mathsf{present}(k')$=false$)\ \land\ \forall k'.\ \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$ )
    y = EMPTY;
  else
    choose k1 $\in$ $\<Ops>$ with $\mathsf{present}$(k1)$\land\neg \exists k'.\ \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$;
    $\mathsf{present}$(k1) = false;
    y = $\mathsf{arg}$(k1);
}
\end{lstlisting}
\end{minipage}
\begin{minipage}[t]{5cm}
\vspace{-3mm}
\begin{lstlisting}
rule ret(enq,k) = {
  $\mathsf{pending}$(k) = false;
  return;
}
rule inv(deq,k) = {
  skip;
}
rule ret(deq,y,k) = {
  return y;
}
\end{lstlisting}
\end{minipage}
%	\caption{The abstract queue implementation $AbsQ$.}
%	\label{fig:signatures}
%\end{figure}

>>>>

%\begin{figure} [t]
\begin{figure}[t]
\hspace{-10mm}
\begin{minipage}[t]{7.5cm}
\scriptsize{
{	%\centering
	\begin{align*}
%		& \mathsf{Operations}, \mathsf{Values} : \mathsf{Sort} \\
		& \text{\underline{Abstract State}} \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{present} : \<Ops> \rightarrow \mathbb{B} \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{pending} : \<Ops> \rightarrow \mathbb{B} \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{arg} : \<Ops> \rightarrow \<Vals> \\
		& \text{{\ttfamily \bfseries function}}\ \mathsf{before} : \<Ops> \times \<Ops> \rightarrow \mathbb{B} \\
%		& \mathsf{stored}, \mathsf{pending} : \mathsf{Operations} &\mbox{ // the status of an enqueue} \\
%		& \mathsf{arg} : \mathsf{Operations} \rightarrow \mathsf{Values} &\mbox{ // the argument of an enqueue} \\
	\end{align*}
}}
\end{minipage}
\begin{minipage}[t]{7.5cm}
\begin{lstlisting}
void enq($\<Vals>$ x):
  atomic rule inv(enq,x,k = new)
  atomic rule ret(enq,k)

$\<Vals>$ deq():
  atomic rule inv(deq,k = new)
  atomic rule lin(deq,y,k)
  atomic rule ret(deq,y,k)

\end{lstlisting}
\end{minipage}

\begin{minipage}[t]{7.5cm}
\vspace{-3mm}
\begin{lstlisting}
rule inv(enq,x,k):
  $\mathsf{present}$(k) := true
  $\mathsf{pending}$(k) := true
  $\mathsf{arg}$(k) := x
  forall k1 with $\mathsf{present}$(k1)$\land \neg \mathsf{pending}$(k1):
    $\mathsf{before}$(k1,k) := true

rule lin(deq,y,k):
  choose a $\in$ $\mathbb{B}$
  if $($a$\vee\forall k'. \mathsf{present}(k')$=false$) \land \forall k'. \mathsf{present}(k')\Rightarrow \mathsf{pending}(k')$:
    y := EMPTY
  else:
    choose k1 $\in$ $\<Ops>$ with $\mathsf{present}$(k1)$\land\neg \exists k'. \mathsf{present}(k')\land \mathsf{before}(k',$k1$)$
    $\mathsf{present}$(k1) := false
    y := $\mathsf{arg}$(k1)

\end{lstlisting}
\end{minipage}
\begin{minipage}[t]{5cm}
\vspace{-3mm}
\begin{lstlisting}
rule ret(enq,k):
  $\mathsf{pending}$(k) := false

rule inv(deq,k):
  skip

rule ret(deq,y,k):
  return y

\end{lstlisting}
\end{minipage}
%	\caption{The abstract queue implementation $AbsQ$.}
%	\label{fig:signatures}
%\end{figure}
